---
title: "final_project"
author: "Anisha Babu"
date: "2/21/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=FALSE, message=FALSE}
# import packages
library(rio)
library(tidyverse)
library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap) # need to cite!!
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(gganimate)
library(gifski)
library(png)

# IF YOU DO NOT HAVE TREE DATA YET
tree_census <- read_csv("https://data.cityofnewyork.us/api/views/uvpi-gqnh/rows.csv")

# IF YOU ALREADY DOWNLOADED TREE DATA
# tree_census <- import("2015_Street_Tree_Census_-_Tree_Data.csv")

# Download NYC neighborhood data 
r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)
nyc_neighborhoods_df <- tidy(nyc_neighborhoods)

```


```{r NYC map, warning=FALSE, message=FALSE}
# Count neighborhood trees
trees_by_neighborhood <- tree_census %>%
  group_by(neighborhood = nta_name) %>%
  summarize(num_trees=n())

# Join data
tree_neighborhood <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., trees_by_neighborhood, by=c("id"="neighborhood")) 

# Tree density heat map
ggplot() + 
  geom_polygon(data=tree_neighborhood, 
               aes(x=long, y=lat, group=group, fill=num_trees),
               color="black") +
  theme_void() +
  labs(title = "Tree Density in NYC",
       fill = "Number of Trees") 

# NOTE: neighborhood names do not match perfectly, so some empty neighborhoods
# CONSIDER: may change to a dot density map

#VV = This visualization 1 definitely works; suggestion - Is there a way to label what the grey areas indicate? The legend does not indicate what the grey stands for, perhaps using another color scheme - green to blue may be? 
```

```{r trunk diameter, warning=FALSE, message=FALSE}
# change values to 0/1 numeric

# This certainly works but tidyverse might be helpful to make this eeven more clear

# tree_census %>% 
#   mutate(tree_census, recode("No" = 0,
#                              "Yes" = 1)) 

tree_census[tree_census == "No"] <- "0"  
tree_census[tree_census == "Yes"] <- "1"


# V = Apart from using tidyverse
# tree_census %>% 
#   mutate(root_stone = numeric(root_stone),
#          root_grate = numeric(root_grate)) #adding other variables

# I might also consider building a function and deploying it
# # make_numeric(variable){
#     variable = as.numeric(variable)
# }
# Then use a map function or a for loop for all the variables
# 1:10 %>%
#df %>% # df consists of all the vatiables needed 
# map(~ as.numeric(.x))
  
tree_census$root_stone <- as.numeric(tree_census$root_stone)
tree_census$root_grate <- as.numeric(tree_census$root_grate)
tree_census$root_other <-as.numeric(tree_census$root_other)
tree_census$trunk_wire <- as.numeric(tree_census$trunk_wire)
tree_census$trnk_light <- as.numeric(tree_census$trnk_light)
tree_census$trnk_other <- as.numeric(tree_census$trnk_other)
tree_census$brch_shoe <- as.numeric(tree_census$brch_shoe)
tree_census$brch_light <- as.numeric(tree_census$brch_light)
tree_census$brch_other <- as.numeric(tree_census$brch_other)

tree_damage <- tree_census %>%
  filter(problems != "None") %>%
  pivot_longer(cols = root_stone:brch_other, 
               names_to = "problem_type", 
               values_to = "count") %>%
  group_by(problem_type) %>%
  filter(count == 1) %>%
  summarize(average_dbh = mean(tree_dbh))
tree_damage$problem_type <- as.factor(tree_damage$problem_type) 
tree_damage$problem_type <- recode(tree_damage$problem_type, 
                                   brch_light="Lights in Branches",
                                   brch_shoe = "Shoes in Branches",
                                   brch_other = "Other Branch Problem",
                                   trunk_wire = "Wire on Trunk",
                                   trnk_light = "Lights on Trunk",
                                   trnk_other = "Other Trunk Problems",
                                   root_stone = "Paving Stones on Root",
                                   root_grate = "Grate on Root",
                                   root_other = "Other Root Problems")

tree_damage %>%
  ggplot(aes(x = problem_type, y = average_dbh)) +
  geom_bar(stat = "identity", aes(fill = problem_type)) +
  labs(title = "Tree Diameter by Problem Type",
       x = "Problem Type",
       y = "Average Tree Diameter") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, hjust=1))

#VV = bar chart is a great choice! 
# Suggestion - may be use a different color scheme, unless you intend to use colors that indicate a continuum; may be use coord_flip() - not sure about this

```

```{r species type, warning=FALSE, message=FALSE}
tree_species <- tree_census %>%
  group_by(nta_name) %>%
  summarize(num_trees=n(), num_species = length(unique(spc_common))) %>%
  pivot_longer(cols = num_trees:num_species, 
               names_to = "type_count",
               values_to = "count")

species_plot <- tree_species %>%
  filter(type_count == "num_species") %>%
  ggplot(aes(x = count, y = type_count)) +
  geom_bar(stat = "identity", fill = "skyblue3", color = "gray70") +
  labs(x = "Tree Species Count") +
  theme(axis.text.y=element_blank(),
        axis.title.y=element_blank())
species_plot <- species_plot +
  transition_states(nta_name,
                    transition_length = 2,
                    state_length = 1) +
  labs(title = 'Number of Tree Species in: {closest_state}')

animate(species_plot,fps = 1)

#VV = My suggestion would be reducing the width of the bar; I am not sure but may be there is a way to modify the time for the switch between categories 

```

# Peer Review - Vinita Vader

## At least three areas of strength
- Great use of left_join
- The playing around with the theme() makes a significant difference in the graph for the better
- Great use of gganimate package
## At least one thing you learned from reviewing their script
- gganimate is very easy to use and can be quite effective

## at least one and no more than three areas for improvement for each visualization.

### Viz 1 - This visualization 1 definitely works; suggestion - Is there a way to label what the grey areas indicate? The legend does not indicate what the grey stands for, perhaps using another color scheme - green to blue may be? 

### Viz 2 - 
Apart from using tidyverse
 tree_census %>% 
   mutate(root_stone = numeric(root_stone),
          root_grate = numeric(root_grate)) #adding other variables

 I might also consider building a function and deploying it
  make_numeric(variable){
     variable = as.numeric(variable)
 }
 Then use a map function or a for loop for all the variables
 1:10 %>%
df %>% # df consists of all the vatiables needed 
 map(~ as.numeric(.x))
 
 
 bar chart is a great choice! 
Suggestion - may be use a different color scheme, unless you intend to use colors that indicate a continuum; may be use coord_flip() - not sure about this

 ### Viz 3 -  My suggestion would be reducing the width of the bar; I am not sure but may be there is a way to modify the time for the switch between categories 
 
## Comments on both the code leading up to (and including) the visualization, and the visualization itself (aesthetics, best practices, etc.).
 - I can see a possible story that can be told with these visualizations
 - The coding is very clear although commenting on certain parts may be helpful
 - I picked up a couple of new packages like gifski and maptools. 
 







